Open Clusters Characterization in Gaia DR2 Using ML Algorithms

### 2020-12-03 thesis/thesis.tex
doc: Update README.md

In this work, we develope a method to characterize those clusters by using \emph{artificial intelligence} tools like clustering
by \emph{K-Means}, or clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering}
model. We are using Gaia DR2 database as data source for testing our models.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.pdf}
      \caption{Proper motions}
      \label{fig:pm_melotte_22}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.pdf}
      \caption{Parallax}
      \label{fig:pm_vec_melotte_22}
    \end{subfigure}
  \end{subfigure}
  \caption{Open Cluster Melotte 22 (Messier 45)}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_6494/pm_ngc_6494.pdf}
      \caption{Proper motions}
      \label{fig:pm_ngc_6494}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_6494/parallax_ngc_6494.pdf}
      \caption{Parallax}
      \label{fig:parallax_ngc_6494}
    \end{subfigure}
  \end{subfigure}
  \caption{Open Cluster NGC 6494}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/raw_pm_melotte_22.pdf}
  \end{subfigure}
  \caption{Melotte 22 proper motions}
  \label{fig:raw_pm_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_parallax_melotte_22.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_hr_diagram_melotte_22.pdf}
    \end{subfigure}
  \end{subfigure}
  \caption{Melotte 22 parallax and H-R diagram}
  \label{fig:raw_parallax_hr_diagram_mellote_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/openclust_catalogue.pdf}
  \caption{OpenClust Catalogue Distribution}
  \label{fig:OpenClustComplete}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/cluster_selection_tier1.pdf}
  \caption{OpenClust Catalogue Selection Distribution}
  \label{fig:OpenClustSelection}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/features_melotte_22.pdf}
  \end{subfigure}
  \caption{Melotte 22 feature pairplot}
  \label{fig:features_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.pdf}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.pdf}
      \caption{N clusters = 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.pdf}
      \caption{N clusters = 5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.pdf}
      \caption{N clusters = 8}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means comparisons with Melotte 22}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.pdf}
      \caption{Proper Motion}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.pdf}
      \caption{Parallax}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means model applied to Melotte 22}
  \label{fig:kmeans_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.pdf}
  \end{subfigure}
  \caption{Melotte 22 H-R diagram with K-Means characterization}
  \label{fig:kmeans_hr_diagram_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.pdf}
      \caption{Proper Motion}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.pdf}
      \caption{Parallax}
    \end{subfigure}
  \end{subfigure}
  \caption{DEC model applied to Melotte 22}
  \label{fig:dec_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.pdf}
  \end{subfigure}
  \caption{Melotte 22 H-R diagram with DEC characterization}
  \label{fig:dec_hr_diagram_melotte_22}
\end{figure}

>>>>

In this work, we develope a method to characterize those clusters by using \emph{artificial intelligence} tools like clustering
by \emph{K-Means}, or clustering based on \emph{Artificial Neural Networks} by implementing the \emph{Deep Embedded Clustering}
model. We are using \emph{Gaia DR2 database} as data source for testing our models.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/pm_melotte_22.png}
      \caption{Proper motions}
      \label{fig:pm_melotte_22}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/parallax_melotte_22.png}
      \caption{Parallax}
      \label{fig:pm_vec_melotte_22}
    \end{subfigure}
  \end{subfigure}
  \caption{Open Cluster Melotte 22 (Messier 45)}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_6494/pm_ngc_6494.png}
      \caption{Proper motions}
      \label{fig:pm_ngc_6494}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_6494/parallax_ngc_6494.png}
      \caption{Parallax}
      \label{fig:parallax_ngc_6494}
    \end{subfigure}
  \end{subfigure}
  \caption{Open Cluster NGC 6494}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/raw_pm_melotte_22.png}
  \end{subfigure}
  \caption{Melotte 22 proper motions}
  \label{fig:raw_pm_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_parallax_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/raw_hr_diagram_melotte_22.png}
    \end{subfigure}
  \end{subfigure}
  \caption{Melotte 22 parallax and H-R diagram}
  \label{fig:raw_parallax_hr_diagram_mellote_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/openclust_catalogue.png}
  \caption{OpenClust Catalogue Distribution}
  \label{fig:OpenClustComplete}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{../figures/cluster_selection_tier1.png}
  \caption{OpenClust Catalogue Selection Distribution}
  \label{fig:OpenClustSelection}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/features_melotte_22.png}
  \end{subfigure}
  \caption{Melotte 22 feature pairplot}
  \label{fig:features_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
      \caption{N clusters = 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
      \caption{N clusters = 5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
      \caption{N clusters = 8}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means comparisons with Melotte 22}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_pm_melotte_22.png}
      \caption{Proper Motion}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_parallax_melotte_22.png}
      \caption{Parallax}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means model applied to Melotte 22}
  \label{fig:kmeans_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/kmeans_hr_diagram_melotte_22.png}
  \end{subfigure}
  \caption{Melotte 22 H-R diagram with K-Means characterization}
  \label{fig:kmeans_hr_diagram_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_pm_melotte_22.png}
      \caption{Proper Motion}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/melotte_22/dec_parallax_melotte_22.png}
      \caption{Parallax}
    \end{subfigure}
  \end{subfigure}
  \caption{DEC model applied to Melotte 22}
  \label{fig:dec_melotte_22}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/melotte_22/dec_hr_diagram_melotte_22.png}
  \end{subfigure}
  \caption{Melotte 22 H-R diagram with DEC characterization}
  \label{fig:dec_hr_diagram_melotte_22}
\end{figure}
